Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-21 paper-ieee-trans.tex

\IEEEPARstart{A}{n}alyzing the European Commission-funded software projects in the last decade we have observed a continuous
increase of the prominence and robustness of testing and validation procedures. The first of many reasons is related to the evolution of the software engineering
techniques (e.g. rise of DevOps practices \cite{zhu}) together with the parallel advancements in the ICT field with virtualization and container
technologies. Another one is the emergence of automation and event-response tools.

Software developers involved in these projects over time have learned  how to make their final products more reliable in order to satisfy requirements of different communities, such as physics.

In this paper, we introduce a set of European Commission (EC)-funded projects that goes from
2001 up to now, where the authors of this summary were and are currently
involved. For each project, we explain how they addressed the challenge in
software reliability. Furthermore, we detail new trends in software reliability.

The reminder of this paper is organized as follows. Section \ref{sec:ev} details the evolution of research on software reliability in EC-projects. Section \ref{sec:ntsr} provides information about new research trends in software reliability. Finally, Section \ref{sec:con} concludes.

>>>>

\IEEEPARstart{W}hen e-Infrastructures in Europe were first conceived, new
services to access distributed resources -- data, computing, instrumental --
were initially considered. Major investments were then injected in order to
develop distributed solutions that provided a seamless and transparent access
to the resources, available for the researchers. One clear example is Worldwide
Large Hadron Collider Computing Grid (WLCG) e-Infrastructure, powered by the
Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing all the data coming from the Large Hadron
Collider particle accelerator.

Years of major developments followed, mainly focused on providing the new
functionalities, hindering the adoption of software engineering methodologies.
The software delivered was by then little tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was deeply compromised, affecting
negatively their exploitation by the scientific communities.

Soon it become apparent that a balance was needed between new features addition
and the reliability of the software produced. Thus, a substantial part of the
software lifecycle should be devoted to deal with the quality of software being
delivered. At this point, evolving software engineering practices were
considered and gradually adopted to define the quality procedures, team
organization and pilot testbeds.

Analyzing the European Commission-funded software projects in the last decade
we have observed a continuous increase of the prominence and robustness of
testing and validation procedures. The first of many reasons is related to the
evolution of the software engineering techniques (e.g. creation of the Agile
Manifesto \cite{agile-manifesto}, rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the ICT field with virtualization and
container technologies. Another one is the emergence of automation and
event-response tools.

In this paper, we introduce a set of European Commission (EC)-funded projects
that goes from 2001 up to now, where the authors of this summary were involved.
For each project, we describe the challenges arised, the methodologies used and
the achievements towards software reliability. The reminder of this paper is
organized as follows. Section \ref{sec:ev} details the evolution of research on
software reliability in EC-projects. Section \ref{sec:ntsr} provides
information about new research trends in software reliability. Finally,
Section \ref{sec:con} concludes.
